---
title: "title"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output:
  github_document: default
  html_notebook: default
---

## Introduction
This notebook creates the plots for ... for my sections of SOC 1120.

## Packages
This notebook requires the following packages:

```{r load-dependencies}
# tidyverse packages
library(dplyr)          # data wrangling
library(ggplot2)        # create plots
library(readr)          # tabular files

# plotting packages
library(RColorBrewer)   # color schemes
library(scales)         # plotting tools

# other packages
library(here)           # manage file paths
```

Two functions from the `source/` directory are also needed:

```{r load-functions}
# plot theme
source(here("source", "cp_sequoiaTheme.R"))

# save plots
source(here("source", "cp_plotSave.R"))
```

The `ggthemes` package is a dependency for `cp_sequoiaTheme()` but it does not need to be loaded - as long as it is installed, the function will work as designed.

## Load Data
All of the raw data for these plots were obtained from ...

They are in separate `.csv` files that we'll load individually:

```{r load-data}

```

## Clean Data
First, we'll tidy up the data so that it is ready for visualization.
