---
title: "title"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output:
  github_document: default
  html_notebook: default
---

## Introduction
This notebook creates the plots for ... for my sections of SOC 1120.

## Packages
This notebook requires the following packages:

```{r load-dependencies}
# tidyverse packages
library(dplyr)          # data wrangling
library(ggplot2)        # create plots

# spatial packages
library(sf)             # spatial data tools
library(albersusa)      # composite projection for US

# other packages
library(ggthemes)       # themes for ggplot2
library(here)           # manage file paths
library(RColorBrewer)   # color ramps
```

Two functions from the `source/` directory are also needed:

```{r load-functions}
# plot theme
source(here("source", "cp_sequoiaTheme.R"))

# save plots
source(here("source", "cp_plotSave.R"))
```

## Load and Clean Data
The data needed for this map are stored in the `albersusa` package. We'll immediately subset and modify the data to simplify what is imported:

```{r load-data}
strokeBelt <- usa_sf(proj = "laea") %>%
  select(iso_3166_2) %>%
  rename(state = iso_3166_2) %>%
  mutate(state = as.character(state))
```

There are three competing definitions that we'll map from [Glymour et al 2007](https://www.ahajournals.org/doi/full/10.1161/STROKEAHA.107.482059), [Karp et al 2016](https://www.ahajournals.org/doi/full/10.1161/STROKEAHA.116.012997), and from the pNational Heart, Lung, and Blood Institute](https://web.archive.org/web/20081001162318/http://www.nhlbi.nih.gov/health/prof/heart/other/sb_spec.pdf):

```{r create-states}
sb_gly07 <- c("AL", "AR", "GA", "MS", "NC", "SC", "TN")
sb_karp16 <- c("AL", "AR", "GA", "LA", "MS", "NC", "SC", "TN")
sb_nhlbi <- c("AL", "AR", "GA", "IN", "KY", "LA", "MS", "NC", "SC", "TN", "VA")
```

We'll take these three definitions and create objects from them:

```{r add-states}
strokeBelt <- strokeBelt %>%
  mutate(
    gly07 = ifelse(state %in% sb_gly07, 1, 0),
    karp16 = ifelse(state %in% sb_karp16, 1, 0),
    nhlbi = ifelse(state %in% sb_nhlbi, 1, 0)
  ) %>%
  mutate(sum = as.factor(gly07+karp16+nhlbi))
```

## Create Stroke Belt Map

```{r}
# define colors
pal <- brewer.pal(n = 4, name = "Purples")
colors <- c("0" = "#bfbfbf", "1" = pal[2], "2" = pal[3], "3" = pal[4])

# generate plot
ggplot() +
  geom_sf(data = strokeBelt, mapping = aes(fill = sum), show.legend = FALSE) +
  geom_sf(data = strokeBelt, fill = NA, color = "#000000") +
  scale_fill_manual(values = colors) +
  cp_sequoiaTheme(background = "transparent", map = TRUE)

# save plot
cp_plotSave(filename = here("results", "strokeBelt.png"), preset = "sm", dpi = 500)
```


